%%%%%%%% TO COMPILE THE DOC, RUN BIBTEX ON EACH SEPARATE AUXILIARY FILE: (because different refsections in the document).

%% bibtex Thesis1-blx.aux
%% bibtex Thesis2-blx.aux
%% pdflatex Thesis

%%%%%%%% DO THIS TWICE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% PhD word limit = 100,000 words

%%%%% Notes: to do

%% create separate section for collaborations and authorship and contribution

%% general introduction to write for end of January

%% Chapter 3: 
    %% relecture
	%% add Figure reproduced from "..." in published article figures
	%% fix figure 3.4
	%% check table of contents / list of tables, figures.

%% TODO chapter 2
	%% relecture
	%% SI: 2.1 => taxonomic corrections: check numbers; text + refs + bibliography
	%% check table of contents / list of tables, figures.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt]{report}
\usepackage{graphicx}
\usepackage[margin = 2.5cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage{setspace}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{lscape}
\usepackage{mdframed}
\usepackage{xcolor}

% checkmarks
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 

% fancy chapter headers
\usepackage{titlesec, blindtext, color}
\definecolor{gray75}{gray}{0.75}
\newcommand{\hsp}{\hspace{20pt}}
\titleformat{\chapter}[hang]{\Huge\bfseries}{\thechapter\hsp\textcolor{gray75}{|}\hsp}{0pt}{\Huge\bfseries}

% captions
\usepackage[labelfont=bf]{caption}
\captionsetup{skip=7pt, font=footnotesize} %,font={stretch=1}

% tables
\usepackage{multirow}
%\usepackage{longtable} 
\usepackage{array}
%\usepackage{verbatim} 
\usepackage{csvsimple}

% Bibliography
\usepackage[backend=bibtex, natbib=true, url=false, doi=true, isbn=false, style=authoryear-comp, eprint=false, maxbibnames= 20, firstinits=true, maxcitenames=2]{biblatex}
\addbibresource{LIBRARY.bib}
\input{biblatex_options.tex}

\setlength\bibitemsep{1.5\itemsep}
% Numbering biblio
\usepackage[numbib,nottoc]{tocbibind}

% For tables from csv
\usepackage{csvsimple}


%Line numbering
\usepackage{lineno}
%\linenumbers

\usepackage{moreverb} % for verbatim ouput

%\usepackage{lmodern}
\usepackage{times}

% line spacing
\renewcommand{\baselinestretch}{1.5}

% spacing for contents (between number and caption)
\usepackage{tocloft}
\setlength{\cftfignumwidth}{3em} % change Xem to suit
\setlength{\cfttabnumwidth}{3em} % change Xem to suit
\setlength{\cftsecnumwidth}{3em}

% clinkable links 
\usepackage[hidelinks]{hyperref}
\hypersetup{
    colorlinks=false, %set true if you want colored links
    linktoc=all     %set to all if you want both sections and subsections linked
    %linkcolor=blue,  %choose some color if you want links to stand out
}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

\input{Title_page}

%\maketitle

% Declarations and acknowledgements here
\chapter*{Declaration}
%I, Adrienne Etard, confirm that the work presented in this thesis is my own. Where information has been derived from other sources, I confirm that this has been indicated in the thesis.
The work presented in this thesis is my own. Some parts have been conducted in collaboration
with other researchers, and this is indicated where relevant. All else is appropriately referenced.


\vspace{3cm}
This document was compiled with {\LaTeX}; the source code and files are available at: \url{https://github.com/AdrienneEtard}

\chapter*{Acknowledgments}

\chapter*{Abstract} %% No more than 300 words
%This report summarises the work I have achieved throughout the first year of my PhD thesis at the Center for Biodiversity and Environment Research at UCL. Many experts state we have entered a new geological epoch, the Anthropocene, characterised by increasing impacts of human activities on the Earth's systems. Since the industrial revolution in the 18$^{\text{th}}$ century, anthropogenic activities have triggered global changes of planetary climate, and have led to the destruction and alteration of the natural habitat of many species. 

%The impacts on global biodiversity have been dramatic. Vertebrate populations have globally declined by 60\% since 1970. Conserving biodiversity is not just a moral obligation. Biodiversity sustain important ecological processes that participate in making the planet habitable by humans. In order to put into place efficient conservation measures, it is vital to understand how different species cope with anthropogenic disturbances. Particularly, understanding what makes species sensitive to anthropogenic threats is valuable to conservation planning.

%In the first Chapter, I present how trait-based approaches can be used to understand how species respond to disturbances, and how anthropogenic changes alter the composition of vertebrate communities. After exposing the gaps in our current understanding, I briefly detail the questions I tackled in this report. The second Chapter focuses on the collection of trait data across terrestrial vertebrates. These data were used in the third Chapter, which investigates how global land-use change impacts the functional diversity of local vertebrate communities. Finally, the last Chapter presents research questions I aim to tackle in the future years of my PhD.

\chapter*{Impact statement}%% No more than 500 words

\chapter*{Data and code access}


%\vspace{1cm}
%\textbf{Key words:} terrestrial vertebrates; traits; land-use change; land-use type; land-use intensity; climate change; functional diversity; vulnerability; metabolic rates.


% CONTENTS
\clearpage
\tableofcontents

\clearpage
%\cleardoublepage
%\addcontentsline{toc}{section}{List of Tables}
\listoftables
%\clearpage

\clearpage
%\addcontentsline{toc}{section}{List of Figures}
\listoffigures

% \clearpage
%List of abbreviations
%\addcontentsline{toc}{section}{List of abbreviations}

%\chapter*{List of abbreviations}
%\input{Abbreviations}

\begin{refsection}

\clearpage
%\addcontentsline{toc}{chapter}{Introduction}
\chapter{General introduction}
%\input{Chapters/Introduction}


\section*{Thesis outline}

Chapter 2: this chapter was conducted in collaboration with 
In this chapter, I collected trait data for terrestrial vertebrates, targeting seven traits that are commonly used in the field (e.g., body mass, trophic level, etc). I assessed the availability of the trait data across vertebrate classes and I investigated whether the data present taxonomic, phylogenetic and spatial biases. This  (the content of this chapter has been published: https://onlinelibrary.wiley.com/doi/full/10.1111/geb.13184).

Chapter 2: Here, I combined the trait data collected in the previous chapter with the PREDICTS database (a database containing records of species occurrence in different land uses), in order to investigate the effects of land-use disturbance on local functional diversity indices (namely functional richness and functional dispersion) with a “space-for-time” approach. The content of this chapter has been accepted for publication. 

Chapter 3: In this chapter, I investigated which traits shape species responses to land-use and climate change. I used the trait data collected in the first chapter. For the land-use aspects, I used the PREDICTS database to investigate whether traits influence species persistence in disturbed land uses. For the climate change aspects, I quantified sensitivity to climate change across vertebrates from the properties of species climatic niche space, and I then investigated whether traits influenced sensitivity to climate change. This chapter is in progress.

Chapter 4: In this chapter, I collected resting metabolic rates for vertebrate species. I combined these metabolic rates with the PREDICTS database to investigate the impacts of land-use disturbance on assemblage-level energetic requirements. I also investigated whether metabolic rates influenced species responses to land-use disturbance, asking whether resting metabolic rates can be considered as response traits. The content of this chapter is going to be submitted as a research article in the coming days. 


%% include a thesis outline
\textcolor{blue}{\textbf{The content of Chapter 2 was published in October 2020 in \textit{Global Ecology and Biogeography}, in a research article co-authored by myself, Tim Newbold and Sophie Morrill, who contributed data on XXX \citep{Etard2020}.}}

\chapter{Global gaps and biases in trait data for terrestrial vertebrates}
\addcontentsline{lof}{chapter}{Chapter 2}
\addcontentsline{lot}{chapter}{Chapter 2}
\input{Chapters/Chapter2}

\chapter{Intensive human land uses negatively affect vertebrate functional diversity}
\addcontentsline{lof}{chapter}{Chapter 3}
\input{Chapters/Chapter3}

\chapter{Different traits explain sensitivity to land use}
%\input{Chapters/Chapter4}

\chapter{Traits \& sensitivity to climate change}
%\input{Chapters/Chapter5}

\chapter{General Discussion}

%% some ideas to include:

%% Land-use and climate change: threats considered independentely => but what if they interact? Additive or not? Synergistic? See Pol Capdevilla's work?

\chapter{Conclusion}


\clearpage
\addcontentsline{toc}{chapter}{Bibliography}
\chapter*{Bibliography}
\renewcommand{\baselinestretch}{1}
\printbibliography[heading=none]

\end{refsection}


%%%%%%%%%%%%%%%%%%%%%%%
%% APPENDICES %% Structured by chapters
\begin{refsection}

% line spacing
\renewcommand{\baselinestretch}{1.15}

% figure and section numbering -- start with S for the appendices

\clearpage

\addcontentsline{toc}{chapter}{Appendices}
\chapter*{Appendices}
\clearpage

%% if ignoring figures and tables from there on in the list of tables and list of figures	
%\captionsetup[figure]{list=no}
%\captionsetup[table]{list=no}

\addcontentsline{toc}{chapter}{Appendix 1: Supporting information for Chapter 1}
\chapter*{Appendix 1: Supporting information for Chapter 1}

\addcontentsline{toc}{chapter}{Appendix 2: Supporting information for Chapter 2}
\addcontentsline{lof}{chapter}{Appendix 2}
\addcontentsline{lot}{chapter}{Appendix 2}
\chapter*{Appendix 2: Supporting information for Chapter 2}
\renewcommand{\thefigure}{S2.\arabic{figure}}
\renewcommand{\thetable}{S2.\arabic{table}}
\renewcommand{\thesection}{S2.\arabic{section}}
\input{Supporting/Chapter2/SI_Chapter2}

\addcontentsline{toc}{chapter}{Appendix 3: Supporting information for Chapter 3}
\addcontentsline{lof}{chapter}{Appendix 3}
\addcontentsline{lot}{chapter}{Appendix 3}
\chapter*{Appendix 3: Supporting information for Chapter 3}
\setcounter{section}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\thefigure}{S3.\arabic{figure}}
\renewcommand{\thetable}{S3.\arabic{table}}
\renewcommand{\thesection}{S3.\arabic{section}}
\input{Supporting/Chapter3/SI_Chapter3}


\addcontentsline{toc}{chapter}{Appendix 4: Supporting information for Chapter 4}
\chapter*{Appendix 4: Supporting information for Chapter 4}

\clearpage
\addcontentsline{toc}{chapter}{Bibliography for the Appendices}
\chapter*{Bibliography for the Appendices}
\renewcommand{\baselinestretch}{1}
\printbibliography[heading=none]
\end{refsection}

\end{document}